---
title: "mixRaschTools Vignettes"
author: "Pamela S Trantham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mixRaschTools Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


These vignettes summarize the methods needed to use mixRaschTools to create a plot of item patameterss which incudes multiple classes for mixed Rasch models, and to create a table containing average ability levels by class.


**Plotting Function**

Sometimes it is useful to examine the differences between the item parameters between classes.  This function creates a plot containing the item parameters of each class to allow a quick examination of how items vary by class.
```{r, include=FALSE}
library(knitr)
```

Using the example data included in the package:

```{r, include=FALSE}
load("/Users/pamela/Desktop/EDPS 845/edps-845/Project/threeclass-ex.rda")
load("/Users/pamela/Desktop/EDPS 845/edps-845/Project/fourclass-ex.rda")
mixRasch.plot <- function(x, ...){
  nc<- length(x$LatentClass)
    plottemp <- sapply(x$LatentClass, function(x) x$item.par$delta.i)
    matplot(plottemp, type = "o", main = "Item Parameters by Class", xlab = "Items", ylab = "Parameters", pch = 20, bg = 1:nc)
    abline(h = 0, lty = 2)
    par(mfrow = c(1, 2))
}

```



```{r, fig.keep=TRUE, fig.cap= "Three Class Model", fig.height=3, fig.width= 6}
mixRasch.plot(threeclass_ex)
```



```{r, fig.keep=TRUE, fig.cap= "Four Class Model", fig.height=3, fig.width= 6}
mixRasch.plot(fourclass_ex)
```


**Average Theta Function**

It is also useful to evaluate the different average ability levels obtained by class in a mixed Rasch model.  The avg.theta function returns a table of average ability for each class found in the model.

Using the example models included in the package:
```{r, include=FALSE}

avg.theta<- function(x){
  nc<- length(x$LatentClass)
  outtheta <- sapply(x$LatentClass, function(x) mean(x$person.par$theta))
  return(outtheta)
}
```

```{r}
avg.theta(threeclass_ex)
avg.theta(fourclass_ex)
```








