---
#Pamela Trantham
title: "EDPS 845 Assignment 1"
author: "Tony Albano"
---

## Overview

Demonstrate your mastery of concepts and procedures covered in *An Introduction to R*, chapters 1 through 9. See Appendix A for examples.

Complete this assignment by responding to each question below using your own original code, in addition to answering any questions. Follow Google's [R coding conventions](https://google.github.io/styleguide/Rguide.xml), and use [R markdown](http://rmarkdown.rstudio.com/index.html) formatting for your responses, with example R code written as

```{r}
# Comment within some example code
```

Your file should have an Rmd extension, and you should check that it compiles to HTML using the *Knit* command in RStudio (this runs `knitr::knit()` on the file). Note that you can already *Knit* this file to see how it works.

## Exercises

### Chapter 1

1. Show that you can access the home page for the HTML R help menu, as well as help files for individual functions.
```{r}
# This opens R help
help.start()
```


2. Demonstrate correct and incorrect naming of R objects.
```{r, eval=FALSE}
#Correct naming of object - Example

object<-c(1,2,3)

#Incorrect naming

21object<- c(1,2,3)
```

3. How do you list all available objects in the R workspace? And how do you remove individual objects?
```{r, eval=FALSE}
#listing
ls()
#removing
rm()
```


### Chapter 2

1. Create and index a vector using a logical, integer, negative integer, and character vector.

```{r}
#create a logical vector
#l<- x < 40
#create an integer vector
x<- c(1, 2, 3, 4)
#create a negative integer vector
y <- x[-(1:3)]
#create a character vector
family<- c("Pam", "Ken", "Tom", "Ray")
```


3. Create a simple data frame containing a numeric variable, character variable, and factor with at least 3 levels. Use the data frame to demonstrate how R codes and handles missing data. 

```{r}
#Simple data frame not removing any NA values
myfamily<- data.frame(familynumber= x, names= family, familycolumn=y)
myfamily[myfamily == 1] <- NA
print(myfamily)
```

### Chapter 3

1. Demonstrate how to check and change the modes and attributes of R objects.
```{r}
state <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa",
"qld", "vic", "nsw", "vic", "qld", "qld", "sa", "tas",
"sa", "nt", "wa", "vic", "qld", "nsw", "nsw", "wa",
"sa", "act", "nsw", "vic", "vic", "act")
statef<-factor(state)
incomes <- c(60, 49, 40, 61, 64, 60, 59, 54, 62, 69, 70, 42, 56,
61, 61, 61, 58, 51, 48, 65, 49, 49, 41, 48, 52, 46,
59, 46, 58, 43)
#Check mode of an object
mode(x)
#change the mode of an object
xchar<- as.character(x)
mode(xchar)
#check length of an object
length(y)
#attributes
attributes(statef)
#change attributes
attrchanged<- attributes(statef)<-NULL
attributes(attrchanged)
```

### Chapter 4

1. Complete the exercise on page 17, where you calculate 95% confidence intervals for state mean incomes.

```{r}
#Page 17


incmeans <- tapply(incomes, statef, mean)
stderr <- function(x) sqrt(var(x)/length(x))
incster<-tapply(incomes, statef, stderr)
length(incomes)
n <- table(statef)
cis<-data.frame(mean=incmeans, stderr=incster, state=names(incmeans),
  lower=qt(.025, n-1),
  upper=qt(.025, n-1, lower.tail=FALSE))

```

2. In what order are the levels of ordered factors stored in?

```{r}
#alphabetical order
```

### Chapter 5

1. Create an array with dimensions `c(3, 6, 2)` and use it to demonstrate array indexing.

```{r}
#Create an array with dimensions 3 by 6 by 2
b <- array(1:36, dim=c(3, 6, 2))

```

2. Create a two dimensional array of 100 integers sampled with replacement from the vector 1:20, and use an index array to set the diagonal values to 0.

```{r}
#Create random sample
d<- c(1:20)
randsample<-sample(d, 100, replace=TRUE)
#Create array to set diagonal values to 0
randomarray<- array(data=randsample, dim=c(10, 10))
i<- array(c(1:10, 10:1),dim=c(10,2))
randomarray[i]
randomarray[i]<-0 
print(randomarray)
```

3. Construct the same array using the `array()` function and using a vector with the `dim()` set. Prove they are the same.

```{r}
#Create array using array()
secondarray<- array(data=randsample, dim=c(4, 25))
#create same array using dim()
seconddim<- randsample ; dim(seconddim) <- c(4, 25)
#prove they are the same
print(secondarray)
print(seconddim)
```

4. Use your array from 5.1, and another with the same dimensions, to demonstrate how arithmetic works with arrays.

```{r}
#create second array of same size as array b
c<- array(37:72, dim=c(3, 6, 2))
#demonstrate arithmetic by multiplying outer product
bc<- outer(b, c, "*")
```

5. Show how to clear the dimensions of an array.
```{r}
vec <- c(bc)
```

### Chapter 6

1. Demonstrate the different ways of subsetting a list.

```{r}
#Subsetting a list
#One way to subset a list:

mtcars[[1]]
#Or
mtcars$cyl

```

2. What is the difference between giving multiple lists to `list()` versus `c()`?
```{r}
#Using the 'list()' command allows the use of existing objects to form lists.  Using the 'c()' command combines components of existing lists together to form new lists.
newlist <- list(mtcars, randsample)
newlist2 <- c(mtcars, randsample)
```

3. How do data frames in R relate to lists?

```{r}
# this is a list with the class "data.frame".  They can include restrictions such as having components made up of vectors that all have the same length and row size.  
```

4. What is the purpose of `attach()`?
```{r}
#The 'attach()' function attaches a database to a search path.  This allows the use to refer to the variables by their names rather than as components of a dataframe. For example, if the dataframe 'mtcars' is not attached, you would have to call the cylinder column in the database with 'mtcars$cyl'.  However, if you attached the mtcars data, you could just use 'cyl' to refer to this column.
```

### Chapter 7

1. Demonstrate the different approaches to reading in data from an external file.
```{r}
#This can be done using the following:
readx<- read.table("/Users/pamela/Desktop/EDPS 845/edps-845/a1/selfeff.txt", header=TRUE, row.names = 1)
scanx <- scan("/Users/pamela/Desktop/EDPS 845/edps-845/a1/selfeff.txt", what="character",skip=1,quiet=TRUE)
datax<- data(cars)

```

### Chapter 8

1. Demonstrate the use of distribution functions to simulate data and calculate probabilities and quantiles.
```{r}
#Simulate random data with Poisson dist, lambda = 4
Pdata<-rpois(15, 4)
#Calculate probabilities for t distribution
2*pt(-2.43, df = 13)
#Calculate Quantiles for F(2, 7)
qf(0.01, 2, 7, lower.tail = FALSE)
```

2. Simulate data from a t distribution and check their normality.
```{r}
#Simulate data from t distribution
tdata<- rt(30, df = n-1)
#Examine normality
stem(tdata)
hist(tdata)
shapiro.test(tdata)
ks.test(tdata, "pnorm", mean = mean(tdata), sd = sqrt(var(tdata)))
```

### Chapter 9

1. What is the difference between `if() else` and `ifelse()`?

```{r}
#ifelse() is a vectorized version of the if/else statement and returns a vector the length of the longest argument, whereas the if() else statement evaluates a single logical argment and returns the appropriate value given the statement.

#example ifelse()
ifelse(randomarray,x,y)

#example if()else
f<-0 
if (f< 0) {
   print("Negative number")
} else if (f > 0) {
   print("Positive number")
} else
   print("Zero")
```

2. Demonstrate the use of a `for()` loop.
```{r}
for(i in 1:10) {
  print(rnorm(10))
}
#sapply will perform the same function for each item in vector

sapply(levels(statef), print)
sapply(levels(statef), nchar)

```

